---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Barbara Mu"
date: "02/09/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

```{r}
dn <- dennys
lq <- laquinta
```

### Exercise 1

```{r}
dn_ak <- dn %>%
  filter(state == "AK")
nrow(dn_ak)
```

```{r}
lq_ak <- lq %>%
  filter(state == "AK")
nrow(lq_ak)
```

### Exercise 2
```{r}
m <- nrow(dn_ak)
n <- nrow(lq_ak)
n_pairing <- m*n
```

There are 6 possible pairs we can calculate their distance. 

### Exercise 3

```{r}
dn_lq_ak <- full_join(dn_ak, lq_ak,
  by = "state"
)
dn_lq_ak
```


### Exercise 4

The combined data frame contains 6 observations as we joined them by `state`. 

### Exercise 5

```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) # distance in km
}
```

The function we should use is mutate(), as we need to create new variable 

### Exercise 6

```{r}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
```

### Exercise 7 

```{r}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```

### Exercise 8

```{r}
library(ggplot2)

ggplot(dn_lq_ak_mindist, aes(x = address.x, y = closest, fill = address.x)) +
  geom_col() +
  coord_flip() +  
  labs(
    title = "Minimum Distance to La Quinta for each Denny's in AK",
    x = "Denny's Address",
    y = "Distance to Nearest La Quinta (km)"
  ) +
  theme_minimal() +
  guides(fill = "none") 
```

### Exercise 9
```{r}
dn_nc <- dn %>% filter(state == "NC")
lq_nc <- lq %>% filter(state == "NC")
dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")

dn_lq_nc <- dn_lq_nc %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))

dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```

### Exercise 10

```{r}
dn_tx <- dn %>% filter(state == "TX")
lq_tx <- lq %>% filter(state == "TX")
dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")

dn_lq_tx <- dn_lq_tx %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))

dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```

### Exercise 11

```{r}
dn_ny <- dn %>% filter(state == "NY")
lq_ny <- lq %>% filter(state == "NY")
dn_lq_ny <- full_join(dn_ny, lq_ny, by = "state")

dn_lq_ny <- dn_lq_ny %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))

dn_lq_ny_mindist <- dn_lq_ny %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```

### Exercise 12

Mitch Hedberg’s joke is most likely to be true in Texas. By checking the distance datasets we created in the last four exercises, I found the smallest 'closet' value is from Texas (the address is 607 Avenue Q, and the distance between the Denny and La Quinta is 0.016)
